The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2012-05-30 smurf_paper.tex
finish draft of cosmology section

\begin{abstract}
  We describe the properties of data from the Submillimetre Common
  User Bolometer Array 2 (SCUBA-2) taken during the Shared Risk
  Observing period from 2010 February 11 to 2010 March 25, and the
  production of maps using the Sub-Millimetre User Reduction Facility
  (SMURF). Our iterative approach is similar to those used to reduce
  data from other ground-based submillimetre telescopes in recent
  years, and recovers angular scales up to the size of the array
  footprint (approximately 2\,arcmin). The method is also able to
  achieve the theoretical white-noise limit of the instrument for
  point-source studies.
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1000\,\micron) have helped revolutionize
several important areas of astrophysics, including: the
characterization of early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, the early stages of planet formation.  With
$\sim10000$ total detectors, SCUBA-2 is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Most submillimetre imaging cameras, to date, have used bolometers to
maximize sensitivities \citep[e.g.][]{griffin2002}. Each pixel
consists of a thermal absorber to collect photons that have passed
through broad-band filters, and the signal is read-out using a
thermometer. The high-frequency broad-band noise in modern systems,
such as SCUBA-2, is limited by photon noise from the instrument and
ambient backgrounds. The low-frequency noise, however, is dominated by
sources such as slow variations in mean backgrounds (such as thermal
variations within the cryostat, and the atmosphere for ground-based
cameras), and drift in the readout electronics. Since the
low-frequency noise is largely correlated between all, or subsets, of
the bolometers in time, it can be suppressed during map-making, since
astronomical signals have the distinct property that they are fixed in
a sky reference frame (assuming they are not time-varying). If
successful, the noise in the resulting map is said to be ``background
noise limited'', meaning that the only significant remaining noise
sources are uncorrelated spatially in the map (the theoretical limit).

There are numerous ways to attack this problem, both in terms of the
data-collection method, and data processing. Historically, both radio
and submillimetre telescope systems have used ``chopping'' to perform
rapid differences between target sky locations and some reference
(e.g., another point on the sky, internal calibrators, and different
frequencies in the case of heterodyne systems). In fact, spatial
chopping was used by SCUBA, the predecessor to SCUBA-2, to remove
low-frequency drifts \citep{holland1999}. However, chopping (in
particular spatial chopping) has the unfortunate side-effect of
filtering out information on scales larger than the chop-throw
\textbf{(references)}.For this reason, modern cameras like SCUBA-2 aim
to provide a high-degree of inherent detector stability, and rapidly
slew the telescope so that astronomical signals up to large scales are
captured in a signal band at considerably higher frequency than the
residual low-frequency drift.

@@

often used various
``chopping'' methods to remove low-frequency drifts -- rapid
differences are made between the sky signal at a given position and an
internal reference signal; the sky signal in two different bands for
heterodyne systems (frequency chopping); and a different position in
the sky \citep[a method employed by SCUBA, the predecessor of
SCUBA-2][]{holland1999}.

@@

\scuba\ is presently unique in terms of the shear volume of data that
it produces on a nightly basis, and the fact that it is observing a
wide variety of science targets for multiple users (unlike dedicated
CMB experiments, for example). For this reason, SMURF has been
designed to use a faster, and less memory-intensive algorithm than
CMB-style maximum likelihood map-makers, although it is less accurate
at reproducing large-scale structures. The basic procedure is to
iteratively model and remove most of the bolometer-bolometer
correlated signal components and independent low-frequency noise, and
then regrid the residual signals. This method is similar in spirit to
the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}. While most of the
analysis described here applies equally to all of the data taken
to-date with \scuba, we note that the SCUBA-2 Shared-Risks Observing
period (S2SRO) which took place from 2010 February 11 to 2010 March 25
used only a single subarray at each of the two observing wavelengths,
and had considerably different noise properties and poorer performance
than the fully-commissioned instrument in 2011.

@@

The \scuba\ bolometer temperatures are read-out using superconducting
transition-edge sensors (TESs), with critical temperatures
$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
in which the transition from a superconductor to a normal material
occurs is extremely small (...typical value...?). If the devices are
biased to a temperature within this transition on large time scales,
then tiny temperature fluctuations on short time-scales will produce
significant changes in the impedence. The entire SCUBA-2 focal plane
is therefore held at a \emph{base temperature} somewhat below
$T_\mathrm{C}$. The TESs then have a small electrical voltage biases,
and individual pixel heaters (using a current bias to apply the
desired Joule heating) are used to place the detectors in the
transition. Only a small detector bias voltage is required since the
heaters compensate for most of the variations in the base temperature,
and other ambient loads (like the atmosphere).  Small temperature (and
therefore impedence) fluctuations produce a changing current through
the TESs. The resulting changing magnetic fields are then amplified
using chains of superconducting quantum interference devices (SQUIDs),
and the larger output current is finally digitized. Each detector has
its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each column
of 40 detectors. Each of the 32 rows is then read-out in sequence at a
row-visit frequency of ...kHz?... This multiplexing scheme greatly
reduces the number of readout and bias lines to each detector, making
such large-format bolometer array designs feasible. It should also be
noted that there is an additional 33rd row of SQUID reasouds that are
not connected to TES. These ``dark SQUIDs'' are used to track
non-thermal noise sources that are common to each column's amplifier
chain.

@@

In order to observe astronomical sources, it is clearly advantageous
that the signals appear in the region of the spectrum that offers the
highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
is achieved by modulating the signal; the original SCUBA primarily
used the JCMT chopping secondary to place sources into the spectrum at
about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
ultimately limited to recovering information on angular scales smaller
than the chop throw (a maximum of a couple of arcmin)... cite
somebody. The primary approach to map-making with SCUBA-2 is instead
to slew the entire telescope at such a rate that angular scales of
interest are crossed on time scales shorter than $1/f_\mathrm{knee}$,
similar to other ground-based instruments (MAMBO, BOLOCAM, SHARC-II,
AzTEC, LABOCA,...). The desire to scan rapidly has two basic
constraints: (i) the telescope cannot physically move faster that
several hundred arcsec\,sec$^{-1}$ without requiring unacceptably
large accelerations to change direction, and (ii) bolometers are
designed with a thermal time constant that balance response with
sensitivity. Taking these constraints into consideration, SCUBA-2 was
ultimately designed to work at a maximum scan velocity of
600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
sample rate of 200\,Hz, which at that scan velocity results in a
sample every 3.0\,arcsec, or approximately one third of the
450\,\micron\ diffraction-limited full-width at half-maximum (FWHM) --
a typical rule-of-thumb for adequately sampling a Gaussian point
spread function (PSF). Similarly, the SCUBA-2 bolometers are designed
to have a time constant of 2 ms \textbf{check that number!} in order
to avoid distorting PSF-sized sources at this maximum scan rate.

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster pattern, or a
series of parallel passes across this field. However, such a scan
yields extremely poor performace for structures that are transverse to
the scan direction, since a long time will have ellapsed between each
parallel step, and the signal will therefore be totally dominated by
the $1/f$ noise. In addition, other noise features may appear in the
data periodically. With such a regularly spaced periodic scan
strategy, these noise features can easily conspire with the scan
pattern to produce similarly periodic noise structures in the map, or
\emph{scan-synchronous} noise. It has therefore been recognized that
scan strategies that visit each region of the map on as many different
time scales, and scan angles as possible offer the best protection
against large-scale noise (...cite some people...); such strategies
are said to offer good \emph{cross-linking}. As a tradeoff between
truly random scan patterns and feasibility of implementation taking
into consideration both mechanical and scheduling constraints, SCUBA-2
presently has three primary scan modes: PONG (cite Borys et al.),
Lissajous, and Daisy... (more citations? Kovacs et al. has a nice
description... also CMB literature).

%-------------------------------------------------
\subsection{Typical bolometer signals}
%-------------------------------------------------

@@

The telescope pointing offsets for this approximately 0.5\,deg
diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
phase of the azimuth offsets from the map centre in this scan pattern
slowly drifts with respect to the elevation offsets, it is clear that
the bolometer and dark squid signals are detecting a noise source that
is correlated only with the azimuthal motion and \emph{not} the
elevation. This behaviour would be expected if if there were a strong
magnetic field fixed with respect to the telescope dome (i.e., the
earth's magnetic field). Since SCUBA-2 is mounted on a Nasmyth
platform, only azimuthal motion will change the effective direction of
such a field with respect to the cryostat. It is also worth noting
that the absolute azimuth of the telescope was close to 180\,deg
(i.e., nearly due-south). We suspect that this observing
configuration, combined with the large amplitude of the scan, may have
conspired to produce these large signals.

@@

One unique feature of the SCUBA-2 data is that one can perform a PCA
analysis of both the 450 and 850\,\micron\ data simultaneously. The
first 6 most significant modes of our analysis are shown in
Fig.~\ref{fig:pca}, with the eigenvectors in the top panels, and maps
of the eigenvalues in each subarray in the bottom panels.

@@

As described in the introduction, rather than pursue a slow but
potentially more accurate \emph{maximum likelihood} method that would
require the inversion of a complex linear problem, we instead use an
iterative approach that estimates and removes most of the correlated
(and in particular low-frequency) noise sources in parallel with a
simplified map estimation. We do note, however, that recent advances
in direct inversion methods such as \citet{patanchon2008} may in fact
be tenable for SCUBA-2 data. A future goal of the software team is to
investigate meshing SMURF (perhaps as a cleaning step and initial
condition) for a SANEPIC-like algorithm.

@@

 We decided not to use the \emph{unbiased} estimator since in
practice it would require accumulating an additional array of values
at every map pixel, and only results in a small difference where there
are less than $\sim10$ samples per pixel (a situation that is almost
never encountered in a SCUBA-2 map except in the edge pixels).

@@

In addition to map estimation, the \model{AST} model can also be used
to perform map-based despiking of the time-series
(Section~\ref{sec:mapdespike}), and to apply constraints to the map to
improve convergences (Section~???).

@@

As in Section~\ref{sec:psd} the FFT of each bolometer time series is
taken, and the PSD calculated. An average white noise level is then
measured from 2 to 10\,Hz, a relatively clean region of the PSD that
tends to lie above the $1/f$ knee, but below the high-frequency line
features for typical bolometer data (Fig.~\ref{fig:pspec}). Taking
this constant level for $\mathbf{P}(f)$ we then calculate the expected
variance of the time-series (in approximately 200\,Hz samples) using
Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
uncorrelated thermal sources (i.e. no other long-scale drifts), with
no high-frequency line features, and without the attenuation at even
higher frequencies by the anti-aliasing filter, this is the
theoretical noise limit of the detectors. These measured variances are
stored, and then used in subsequent iterations to weight the data
points when calculating the map estimate (Section~\ref{sec:ast}). They
are also used for convergence tests (Section~\ref{sec:converge}).

@@

The map-maker will halt if the convergence criteria have been
met. Presently there are two numerical tests that may be performed by
SMURF, or alternatively, a fixed number of iterations may be
specified. The first numerical test is an approximate check of the
change in reduced chi-squared, $\chi^2_\mathrm{r}$. The standard
deviations, $\sigma_i$, are measured for each of the residual
bolometer signals once the modeled signal components have been
removed. In the early iterations, the residual signals usually contain
both white noise, and other long-timescale features. However, once the
solution has converged, this signal should look approximately
white. Once the white nosie, $\mathbf{n}^\mathrm{w}_i(t)$, has been
measured from the bolometer PSDs in \model{NOI}
(Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated as $(1/N)
\sum_i \sigma^2_i / \mathbf{n}^\mathrm{w}_i(t)$, where $i$ runs over
the $N$ bolometers. In other words, it is the average ratio between
the measured bolometer variances and their theoretical white-noise
levels. Note that this expression \emph{does not account for the
degrees of freedom}. Clearly there are a large number of parameters in
the model, which will ``fit-out'' some of the uncorrelated white
noise, and potentially bias this estimate of $\chi^2_r$ low. One
approach would be to keep track of the degrees of freedom, and
therefore provide a correction, as in \citet{kovacs2008}. However,
note that we are only using this value as a convergence test, and its
value is not used in an absolute sense. Furthermore, the reference
white noise values are calculated only once in SMURF
(Section~\ref{sec:noi}), and the number of model parameters are fixed,
so the value of $\chi^2_r$ should decrease monotonically if the
solution is behaving well. The convergence criterion is met if the
change in subsequent values of $\chi^2_r$ is smaller than the
parameter \texttt{chitol}, which is set to $10^{-3}$ by default.

It was immediately found in the analysis of SCUBA-2 data that the
$\chi^2_r$ test described above was not a sufficient stopping
criterion. The reason for this is that various components of the model
are, under normal circumstances, highly degenerate. Once simple
example is the degeneracy between large-scale astronomical structures
(larger than the array footprint), and the common-mode rejection step
(\model{COM}), which will be illustrated in
Section~\ref{sec:point}. Since the flatness of the residual only
reflects the ability of the model to fit the shape of the bolometer
time-series, it does not necessarily correlate with convergence of the
most important model component: the map. Indeed, examination of the
map after large numbers of iterations (e.g., $>10$), shows the
presence of large-scale structures that are anti-correlated with
structures in other model components. For this reason, a second
map-based convergence test was added.

@@

\textit{Surely flatfields are discussed in the main SCUBA-2 paper so
don't really have any useful purpose here since they aren't
technically part of the map-maker.}

@@

The accurate measurement of positions and brightnesses of known point
sources are necessary in real-time to establish telescope pointing
offsets and focus. They are also necessary to measure the FCF
(absolute calibration), and hence noise performance of the instrument
in astronomical units.  In this example we produce a 450\,\micron\ map
of Uranus (observation 26 on 20111017), which is a nearly point-like
source for SCUBA-2 that is commonly used as a primary flux
calibrator. The now standard CV daisy scan pattern was used, with a
scan speed of 155 arcsec\,sec$^{-1}$. We perform several different
reductions of the data to illustrate the purpose of different model
components and the convergence properties of the solution
(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
grid of azimuth (horizontal) and elevation (vertical) offsets from the
position of Uranus (the origin), using 2\,arcsec pixels.

@@

One improvement is to simply restrict the low-frequency signal that
goes into the maps by applying a high-pass filter to the bolometer
time-series. A third reduction of the data uses the default map-making
parameters, as described in Section~\ref{sec:algorithm}, which adds
the \model{FLT} model to accomplish this task immediately prior to map
estimation. We set the filter edge based on an angular scale of
200\,arcsec, which given the scan speed of 155\,arcsec\,sec$^{-1}$,
corresponds to a frequency of 0.78\,Hz. Now using the automatic
map-based convergence test (Section ???), the resulting map converges
after 10 iterations (Fig.~\ref{fig:pointmaps}c). Both the circular
streaks and the large-scale gradient are now removed, but they have
been replaced by an obvious, circularly-symmetric ringing pattern
about Uranus. The reason for this ringing is that the hard-edged
high-pass filter in frequency space is equivalent to a $\sinc$
function-like response in map space. Since the scan pattern is
isotropic (scans equally at all position angles), and there is a
bright point-like source, the result is an azimuthally-symmetric
$\sinc$ function-like pattern in the map about Uranus. Note that due
to the iterative nature of the solution, much of the signal in the
source is actually removed for subsequent iterations of \model{FLT},
but as in the previous case, degeneracies between the map and the
\model{FLT} and \model{COM} signal components can still lead to
complicated structures.

@@

\scuba\ surveys designed to detect extremely faint point-sources
(e.g. high-redshift star-forming galaxies, and debris disks) are
ideally limited by the white-noise performance of the instrument. The
approach described here for maximizing the S/N of point-sources
involves three major steps: (i) generating a map that removes most
low-frequency noise sources with approximately linear response,
without prior knowledge of the location of sources; (ii)
\emph{whitening} (essentially flattening) the map to make the noise
properties of the map easier to deal with; and (iii) detecting point
sources using a \emph{matched-filter}. Note that variations on this
general procedure have been used extensively in the submm cosmology
community using previous instruments
\citep[e.g.][]{scott2002,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
In this section we will reduce scans of the Lockman Hole taken in 2010
during S2SRO as a pilot project for the Cosmology Legacy Survey. It
consists of $\sim8.5$\,hours of data taken in 36 separate scans
(average 15\,min. each) spread over February and March, covering an
area of about 50\,arcmin$^2$. The full list of dates and observation
numbers includes 20100218: 63, 64, 65, 70, 71, 72, 90, 91, 92, 97, 98,
99; 20100220: 111, 112, 113, 118, 128, 129, 130; 20100303: 61, 62, 64,
69, 70, 72, 73, 74, 75; 20100309: 87, 88, 90, 91; and 20100311: 59,
64, 65, 72. These observations represent about 80\% of the total data
taken as part of the project; the remaining observations were dropped
due to problems keeping the arrays properly locked during S2SRO, and
were easily identified by their highly variable and erratic bolometer
time-series (which resulted in maps full of large streaks).

>>>>

\begin{abstract}
  We describe the properties of data from the Submillimetre Common
  User Bolometer Array 2 (\scuba) taken during the Shared Risk
  Observing period from 2010 February 11 to 2010 March 25, and the
  production of maps using the Sub-Millimetre User Reduction Facility
  (SMURF). Our iterative approach is similar to those used to reduce
  data from other ground-based submillimetre telescopes in recent
  years, and recovers angular scales up to the size of the array
  footprint (approximately 2\,arcmin). The method is also able to
  achieve the theoretical white-noise limit of the instrument for
  point-source studies.
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1000\,\micron) have helped revolutionize
several important areas of astrophysics, including: the
characterization of early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, the early stages of planet formation.  With
$\sim10000$ total detectors, \scuba\ is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Most submillimetre imaging cameras, to date, have used bolometers to
maximize sensitivities \citep[e.g.][]{griffin2002}. Each pixel
consists of a thermal absorber to collect photons that have passed
through broad-band filters, and the signal is read-out using a
thermometer. The high-frequency broad-band noise in modern systems,
such as \scuba, is limited by photon noise from the instrument and
ambient backgrounds. The low-frequency noise, however, is dominated by
sources such as slow variations in mean backgrounds (such as thermal
variations within the cryostat, and the atmosphere for ground-based
cameras), and drift in the readout electronics. Since the
low-frequency noise is largely correlated between all, or subsets, of
the bolometers in time, it can be suppressed during map-making, since
astronomical signals have the distinct property that they are fixed in
a sky reference frame (assuming they are not time-varying). If
successful, the noise in the resulting map is said to be ``background
noise limited'', meaning that the only significant remaining noise
sources are uncorrelated spatially in the map (the theoretical limit).

There are numerous ways to attack this problem, both in terms of the
data-collection method, and data processing. Historically, both radio
and submillimetre telescope systems have used ``chopping'' to perform
rapid differences between target sky locations and some reference
(e.g., another point on the sky, internal calibrators, and different
frequencies in the case of heterodyne systems). In fact, spatial
chopping was used by SCUBA, the predecessor to \scuba, to remove
low-frequency drifts \citep{holland1999}. However, chopping (in
particular spatial chopping) has the unfortunate side-effect of
filtering out information on scales larger than the chop-throw
\textbf{(references)}.For this reason, modern cameras like \scuba\ aim
to provide a high-degree of inherent detector stability, and rapidly
slew the telescope so that astronomical signals up to large scales are
captured in a signal band at considerably higher frequency than the
residual low-frequency drift.

@@

often used various
``chopping'' methods to remove low-frequency drifts -- rapid
differences are made between the sky signal at a given position and an
internal reference signal; the sky signal in two different bands for
heterodyne systems (frequency chopping); and a different position in
the sky \citep[a method employed by SCUBA, the predecessor of
\scuba][]{holland1999}.

@@

\scuba\ is presently unique in terms of the shear volume of data that
it produces on a nightly basis, and the fact that it is observing a
wide variety of science targets for multiple users (unlike dedicated
CMB experiments, for example). For this reason, SMURF has been
designed to use a faster, and less memory-intensive algorithm than
CMB-style maximum likelihood map-makers, although it is less accurate
at reproducing large-scale structures. The basic procedure is to
iteratively model and remove most of the bolometer-bolometer
correlated signal components and independent low-frequency noise, and
then regrid the residual signals. This method is similar in spirit to
the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}. While most of the
analysis described here applies equally to all of the data taken
to-date with \scuba, we note that the \scuba\ Shared-Risks Observing
period (S2SRO) which took place from 2010 February 11 to 2010 March 25
used only a single subarray at each of the two observing wavelengths,
and had considerably different noise properties and poorer performance
than the fully-commissioned instrument in 2011.

@@

The \scuba\ bolometer temperatures are read-out using superconducting
transition-edge sensors (TESs), with critical temperatures
$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
in which the transition from a superconductor to a normal material
occurs is extremely small (...typical value...?). If the devices are
biased to a temperature within this transition on large time scales,
then tiny temperature fluctuations on short time-scales will produce
significant changes in the impedence. The entire \scuba\ focal plane
is therefore held at a \emph{base temperature} somewhat below
$T_\mathrm{C}$. The TESs then have a small electrical voltage biases,
and individual pixel heaters (using a current bias to apply the
desired Joule heating) are used to place the detectors in the
transition. Only a small detector bias voltage is required since the
heaters compensate for most of the variations in the base temperature,
and other ambient loads (like the atmosphere).  Small temperature (and
therefore impedence) fluctuations produce a changing current through
the TESs. The resulting changing magnetic fields are then amplified
using chains of superconducting quantum interference devices (SQUIDs),
and the larger output current is finally digitized. Each detector has
its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each column
of 40 detectors. Each of the 32 rows is then read-out in sequence at a
row-visit frequency of ...kHz?... This multiplexing scheme greatly
reduces the number of readout and bias lines to each detector, making
such large-format bolometer array designs feasible. It should also be
noted that there is an additional 33rd row of SQUID reasouds that are
not connected to TES. These ``dark SQUIDs'' are used to track
non-thermal noise sources that are common to each column's amplifier
chain.

@@

In order to observe astronomical sources, it is clearly advantageous
that the signals appear in the region of the spectrum that offers the
highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
is achieved by modulating the signal; the original SCUBA primarily
used the JCMT chopping secondary to place sources into the spectrum at
about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
ultimately limited to recovering information on angular scales smaller
than the chop throw (a maximum of a couple of arcmin)... cite
somebody. The primary approach to map-making with \scuba\ is instead
to slew the entire telescope at such a rate that angular scales of
interest are crossed on time scales shorter than $1/f_\mathrm{knee}$,
similar to other ground-based instruments (MAMBO, BOLOCAM, SHARC-II,
AzTEC, LABOCA,...). The desire to scan rapidly has two basic
constraints: (i) the telescope cannot physically move faster that
several hundred arcsec\,sec$^{-1}$ without requiring unacceptably
large accelerations to change direction, and (ii) bolometers are
designed with a thermal time constant that balance response with
sensitivity. Taking these constraints into consideration, \scuba\ was
ultimately designed to work at a maximum scan velocity of
600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
sample rate of 200\,Hz, which at that scan velocity results in a
sample every 3.0\,arcsec, or approximately one third of the
450\,\micron\ diffraction-limited full-width at half-maximum (FWHM) --
a typical rule-of-thumb for adequately sampling a Gaussian point
spread function (PSF). Similarly, the \scuba\ bolometers are designed
to have a time constant of 2 ms \textbf{check that number!} in order
to avoid distorting PSF-sized sources at this maximum scan rate.

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster pattern, or a
series of parallel passes across this field. However, such a scan
yields extremely poor performace for structures that are transverse to
the scan direction, since a long time will have ellapsed between each
parallel step, and the signal will therefore be totally dominated by
the $1/f$ noise. In addition, other noise features may appear in the
data periodically. With such a regularly spaced periodic scan
strategy, these noise features can easily conspire with the scan
pattern to produce similarly periodic noise structures in the map, or
\emph{scan-synchronous} noise. It has therefore been recognized that
scan strategies that visit each region of the map on as many different
time scales, and scan angles as possible offer the best protection
against large-scale noise (...cite some people...); such strategies
are said to offer good \emph{cross-linking}. As a tradeoff between
truly random scan patterns and feasibility of implementation taking
into consideration both mechanical and scheduling constraints, \scuba
presently has three primary scan modes: PONG (cite Borys et al.),
Lissajous, and Daisy... (more citations? Kovacs et al. has a nice
description... also CMB literature).

%-------------------------------------------------
\subsection{Typical bolometer signals}
\label{sec:bolosignal}
%-------------------------------------------------

@@

The telescope pointing offsets for this approximately 0.5\,deg
diameter scan are also shown in Fig.~\ref{fig:magpickup}. Since the
phase of the azimuth offsets from the map centre in this scan pattern
slowly drifts with respect to the elevation offsets, it is clear that
the bolometer and dark squid signals are detecting a noise source that
is correlated only with the azimuthal motion and \emph{not} the
elevation. This behaviour would be expected if if there were a strong
magnetic field fixed with respect to the telescope dome (i.e., the
earth's magnetic field). Since \scuba\ is mounted on a Nasmyth
platform, only azimuthal motion will change the effective direction of
such a field with respect to the cryostat. It is also worth noting
that the absolute azimuth of the telescope was close to 180\,deg
(i.e., nearly due-south). We suspect that this observing
configuration, combined with the large amplitude of the scan, may have
conspired to produce these large signals.

@@

One unique feature of the \scuba\ data is that one can perform a PCA
analysis of both the 450 and 850\,\micron\ data simultaneously. The
first 6 most significant modes of our analysis are shown in
Fig.~\ref{fig:pca}, with the eigenvectors in the top panels, and maps
of the eigenvalues in each subarray in the bottom panels.

@@

As described in the introduction, rather than pursue a slow but
potentially more accurate \emph{maximum likelihood} method that would
require the inversion of a complex linear problem, we instead use an
iterative approach that estimates and removes most of the correlated
(and in particular low-frequency) noise sources in parallel with a
simplified map estimation. We do note, however, that recent advances
in direct inversion methods such as \citet{patanchon2008} may in fact
be tenable for \scuba\ data. A future goal of the software team is to
investigate meshing SMURF (perhaps as a cleaning step and initial
condition) for a SANEPIC-like algorithm.

@@

 We decided not to use the \emph{unbiased} estimator since in
practice it would require accumulating an additional array of values
at every map pixel, and only results in a small difference where there
are less than $\sim10$ samples per pixel (a situation that is almost
never encountered in a \scuba\ map except in the edge pixels).

@@

In addition to map estimation, the \model{AST} model can also be used
to perform map-based despiking of the time-series
(Section~\ref{sec:mapdespike}), and to apply constraints to the map to
improve convergences (Section~\ref{sec:converge}).

@@

First, the bolometer PSDs are calculated as in Eq.~\ref{eq:psd}. An
average white noise level is then measured from 2 to 10\,Hz, a
relatively clean region of the PSD that tends to lie above the $1/f$
knee, but below the high-frequency line features for typical bolometer
data (Fig.~\ref{fig:pspec}). Taking this constant level for
$\mathbf{P}(f)$ we then calculate the expected variance of the
time-series (in approximately 200\,Hz samples) using
Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
uncorrelated thermal sources (i.e. no other long-scale drifts), with
no high-frequency line features, and without the attenuation at even
higher frequencies by the anti-aliasing filter, this is the
theoretical noise limit of the detectors. These measured variances are
stored, and then used in subsequent iterations to weight the data
points when calculating the map estimate (Section~\ref{sec:ast}). They
are also used for convergence tests (Section~\ref{sec:converge}).

@@

The map-maker will halt if the convergence criteria have been
met. Presently there are two numerical tests that may be performed by
SMURF, or alternatively, a fixed number of iterations may be
specified. The first numerical test is an approximate check of the
change in reduced chi-squared, $\chi^2_\mathrm{r}$. The standard
deviations, $\sigma_i$, are measured for each of the residual
bolometer signals once the modeled signal components have been
removed. In the early iterations, the residual signals usually contain
both white noise, and other long-timescale features. However, once the
solution has converged, this signal should look approximately
white. Once the white nosie, $\mathbf{n}^\mathrm{w}_i(t)$, has been
measured from the bolometer PSDs in \model{NOI}
(Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated as $(1/N)
\sum_i \sigma^2_i / \mathbf{n}^\mathrm{w}_i(t)$, where $i$ runs over
the $N$ bolometers. In other words, it is the average ratio between
the measured time-series bolometer variances and their white-noise
levels measured between 2--10\,Hz. Note that this expression
\emph{does not account for the degrees of freedom}. Clearly there are
a large number of parameters in the model, which will ``fit-out'' some
of the uncorrelated white noise, and potentially bias this estimate of
$\chi^2_r$ low. One approach would be to keep track of the degrees of
freedom, and therefore provide a correction, as in
\citet{kovacs2008}. However, note that we are only using this value as
a convergence test, and its value is not used in an absolute
sense. Furthermore, the reference white noise values are calculated
only once in SMURF (Section~\ref{sec:noi}), and the number of model
parameters are fixed, so the value of $\chi^2_r$ should decrease
monotonically if the solution is behaving well. The convergence
criterion is met if the change in subsequent values of $\chi^2_r$ is
smaller than the parameter \texttt{chitol}, which is set to $10^{-3}$
by default.

It was immediately found in the analysis of \scuba\ data that the
$\chi^2_r$ test described above was not a sufficient stopping
criterion. The reason for this is that various components of the model
are, under normal circumstances, highly degenerate. One simple example
is the degeneracy between large-scale astronomical structures (larger
than the array footprint), and the common-mode rejection step
(\model{COM}), which will be illustrated in
Section~\ref{sec:point}. Since the flatness of the residual only
reflects the ability of the model to fit the shape of the bolometer
time-series, it does not necessarily correlate with convergence of the
most important model component: the map. Indeed, examination of the
map after large numbers of iterations (e.g., $>10$), shows the
presence of large-scale structures that are anti-correlated with
structures in other model components. For this reason, a second
map-based convergence test was added.

@@

\textit{Surely flatfields are discussed in the main \scuba\ paper so
don't really have any useful purpose here since they aren't
technically part of the map-maker.}

@@

The accurate measurement of positions and brightnesses of known point
sources are necessary in real-time to establish telescope pointing
offsets and focus. They are also necessary to measure the FCF
(absolute calibration), and hence noise performance of the instrument
in astronomical units.  In this example we produce a 450\,\micron\ map
of Uranus (observation 26 on 20111017), which is a nearly point-like
source for \scuba\ that is commonly used as a primary flux
calibrator. The now standard CV daisy scan pattern was used, with a
scan speed of 155 arcsec\,sec$^{-1}$. We perform several different
reductions of the data to illustrate the purpose of different model
components and the convergence properties of the solution
(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
grid of azimuth (horizontal) and elevation (vertical) offsets from the
position of Uranus (the origin), using 2\,arcsec pixels.

@@

One improvement is to simply restrict the low-frequency signal that
goes into the maps by applying a high-pass filter to the bolometer
time-series. A third reduction of the data uses the default map-making
parameters, as described in Section~\ref{sec:algorithm}, which adds
the \model{FLT} model to accomplish this task immediately prior to map
estimation. We set the filter edge based on an angular scale of
200\,arcsec, which given the scan speed of 155\,arcsec\,sec$^{-1}$,
corresponds to a frequency of 0.78\,Hz. Now using the automatic
map-based convergence test (Section~\ref{sec:converge}), the resulting
map converges after 10 iterations (Fig.~\ref{fig:pointmaps}c). Both
the circular streaks and the large-scale gradient are now removed, but
they have been replaced by an obvious, circularly-symmetric ringing
pattern about Uranus. The reason for this ringing is that the
hard-edged high-pass filter in frequency space is equivalent to a
$\sinc$ function-like response in map space. Since the scan pattern is
isotropic (scans equally at all position angles), and there is a
bright point-like source, the result is an azimuthally-symmetric
$\sinc$ function-like pattern in the map about Uranus. Note that due
to the iterative nature of the solution, much of the signal in the
source is actually removed for subsequent iterations of \model{FLT},
but as in the previous case, degeneracies between the map and the
\model{FLT} and \model{COM} signal components can still lead to
complicated structures.

@@

\scuba\ surveys designed to detect extremely faint point-sources
(e.g. high-redshift star-forming galaxies, and debris disks) are
ideally limited by the white-noise performance of the instrument. The
approach described here for maximizing the S/N of point-sources
involves three major steps: (i) generating a map that removes most
low-frequency noise sources with approximately linear response,
without prior knowledge of the location of sources; (ii)
\emph{whitening} (essentially flattening) the map to make the noise
properties of the map easier to deal with; and (iii) detecting point
sources using a \emph{matched-filter}. Note that variations on this
general procedure have been used extensively in the submm cosmology
community using previous instruments
\citep[e.g.][]{scott2002,laurent2005,coppin2006,scott2008,perera2008,devlin2009}.
In this section we will reduce scans of the Lockman Hole taken during
S2SRO as a pilot project for the Cosmology Legacy Survey. It consists
of $\sim8.5$\,hours of data taken in 36 separate scans (average
15\,min. each) spread over February and March 2010, covering an area
of about 50\,arcmin$^2$. The full list of dates and observation
numbers includes 20100218: 63, 64, 65, 70, 71, 72, 90, 91, 92, 97, 98,
99; 20100220: 111, 112, 113, 118, 128, 129, 130; 20100303: 61, 62, 64,
69, 70, 72, 73, 74, 75; 20100309: 87, 88, 90, 91; and 20100311: 59,
64, 65, 72. These observations represent about 80\% of the total data
taken as part of the project; the remaining observations were dropped
due to problems keeping the arrays properly locked during S2SRO, and
were easily identified by their highly variable and erratic bolometer
time-series (which resulted in maps full of large streaks).
